---
title: "lesson9_student"
author: "zegildo"
date: "24/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lesson 9
Lições sobre as aulas 9

## Quizz 3

Let's start by examining two variables in the data set. The scatterplot is a powerful tool to help you understand the relationship between two continuous variables.

We can quickly see if the relationship is linear or not. In this case, we can use a variety of diamond
characteristics to help us figure out whether the price advertised for any given diamond is reasonable or a rip-off.

Let's consider the price of a diamond and it's carat weight. Create a scatterplot of price (y) vs carat weight (x). Limit the x-axis and y-axis to omit the top 1% of values.

```{r Q3}
library(ggplot2)
ggplot(data = diamonds, aes(x = carat , y=price)) +
  geom_point()+
  stat_smooth(method = 'lm')+
  xlim(0, quantile(diamonds$carat, 0.99)) + 
  ylim(0, quantile(diamonds$price, 0.99))
```


## Quizz 8
Create two histograms of the price variable and place them side by side on one output image. We’ve put some code below to get you started. The first plot should be a histogram of price and the second plot should transform the price variable using log10.

Set appropriate bin widths for each plot. ggtitle() will add a title to each histogram.


```{r Q8}
library(gridExtra)
library(ggplot2)
diamonds_price <- subset(diamonds, !is.na(price))
plot1 <- qplot(price, data = diamonds_price, binwidth=100) + 
  ggtitle('Price')

plot2 <- qplot(log10(price), data = diamonds_price, binwidth=0.01) +
  ggtitle('Price (log10)')

grid.arrange(plot1,plot2, ncol=2)
```


## Quizz 12
Add a layer to adjust the features of the scatterplot. Set the transparency to one half, the size to three-fourths, and jitter the points.

```{r Q12}
library(scales)
library(ggplot2)
cuberoot_trans <- trans_new(name = "cuberoot",                            transform = function(x) x^(1/3),
                       inverse = function(x) x^3);

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha=1/2, size=3/4, position="jitter") + 
  scale_x_continuous(trans = cuberoot_trans, limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

## Quizz 12

Adjust the code below to color the points by clarity. A layer called scale_color_brewer() has been added to adjust the legend and provide custom colors.

See if you can figure out what it does. Links to resources are in the Instructor Notes.You will need to install the package RColorBrewer in R to get the same colors and color palettes.

```{r Q13}
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color=clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div', palette = 'GnBu',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans, limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

## Quizz 22

Lienar model

```{r Q22}
library(ggplot2)
library(memisc)
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data=diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)

```
## Quizz 23
Adding a new information

```{r Q23}
thisDiamond <- data.frame(carat = 1.00, cut = "Very Good", color = "I", clarity="VS1")

modelEstimate <- predict(m5, newdata = thisDiamond, 
                         interval = "prediction", level =.95)

```


