Lesson 7
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
setwd('C:/Users/usaurio/Documents/Cursos/R/estudos/bases/')
library(dplyr)
pf <- read.delim('pseudo_facebook.tsv')

fc_by_age_gender <- group_by(subset(pf, !is.na(gender)), age, gender)

pf.fc_by_age_gender <- summarise(fc_by_age_gender,
                                 mean_friend_count = mean(friend_count),
                                 median_friend_count = median(friend_count) ,
                                 n = n())
```

***

### Plotting Conditional Summaries
Plotting Conditional Summaries

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = mean_friend_count, color = gender),
       data = pf.fc_by_age_gender) +
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.wide$ratio <- with(pf.fc_by_age_gender.wide, female/male)

ggplot(data = pf.fc_by_age_gender.wide,
       aes(x = age, y = ratio)) + 
  geom_line() + 
  geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + 
  
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure > 0), summary(friend_count/tenure))
```

***

### Friendships Initiated
 Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.

You need to make use of the variables tenure, friendships_initiated, and year_joined.bucket.

You also need to subset the data to only consider user with at least one day of tenure.

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure > 0)) +
  
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) 
  
  #geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)



```

```{r Smooth}
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```
***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
library(ggplot2)
setwd('C:/Users/usaurio/Documents/Cursos/R/estudos/bases/')
yo <- read.csv('yogurt.csv')
yo$id <- as.factor(yo$id)
ggplot(yo, aes(price)) +
  geom_histogram()
```

***

### Number of Purchases
Create a new variable called all.purchases, which gives the total counts of yogurt for each observation or household.

One way to do this is using the transform function. You can look up the function transform and run the examples of code at the bottom of the documentation to figure out what it does.

The transform function produces a data frame so if you use it then save the result to 'yo'!

OR you can figure out another way to create the variable.

```{r Number of Purchases}
yo$all.purcharses <- with(yo, (strawberry + blueberry + pina.colada + plain + mixed.berry))
```

***

### Prices over Time
Create a scatterplot of price vs time.

```{r Prices over Time}
ggplot(aes(x = time, y = price),data = yo) +
  geom_jitter(alpha=1/4, shape = 21, fill=I('#F79420'))

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(800)
sample.ids <- sample(levels(yo$id), 16)

ggplot(data = subset(yo, id %in% sample.ids), 
       aes(x = time, y = price)) +
  facet_wrap( ~ id) +
  geom_line() + 
  geom_point(aes(size = all.purcharses), pch = 1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
setwd('C:/Users/usaurio/Documents/Cursos/R/estudos/bases/')
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

